---
title: "hw5"
author: "Senna"
date: "2024-11-05"
output: github_document
---


```{r}
library(tidyverse)
library(broom)
library(purrr)
```

### Problem 1

Suppose you put n people in a room, and want to know the probability that at least two people share a birthday. For simplicity, we’ll assume there are no leap years (i.e. there are only 365 days) and that birthdays are uniformly distributed over the year (which is actually not the case).

Write a function that, for a fixed group size, randomly draws “birthdays” for each person; checks whether there are duplicate birthdays in the group; and returns TRUE or FALSE based on the result.

Next, run this function 10000 times for each group size between 2 and 50. For each group size, compute the probability that at least two people in the group will share a birthday by averaging across the 10000 simulation runs. Make a plot showing the probability as a function of group size, and comment on your results

```{r}

set.seed(1)

has_duplicate_birthday = function(n) {
  birthdays = sample(1:365, n, replace = TRUE)
  return(length(birthdays) != length(unique(birthdays)))
} #FALSE means there are no duplicate. TRUE means there are duplicate. 


group_sizes = 2:50  
num_it = 10000

p = numeric(length(group_sizes))

for (i in seq_along(group_sizes)) {
  n = group_sizes[i]
  
  sim_results = replicate(num_it, has_duplicate_birthday(n))
  p[i] = mean(sim_results)
}



ggplot(results_df, aes(x = group_sizes, y = p)) +
  geom_line(color = "blue") +
  geom_point(color = "red") +
  labs(title = "Probability of Shared Birthday vs. Group Size",
       x = "Group Size",
       y = "Probability of Shared Birthday") +
  theme_minimal()

```

### Problem 2

When designing an experiment or analysis, a common question is whether it is likely that a true effect will be detected – put differently, whether a false null hypothesis will be rejected. The probability that a false null hypothesis is rejected is referred to as power, and it depends on several factors, including: the sample size; the effect size; and the error variance. In this problem, you will conduct a simulation to explore power in a one-sample t-test.

First set the following design elements:
Fix n=30
Fix σ=5
Set μ=0
Generate 5000 datasets from the model x∼Normal[μ,σ]

```{r}
n = 30
sigma = 5
mu = 0
num_datasets = 5000

set.seed(123)  
datasets <- vector("list", num_datasets) 

for (i in 1:num_datasets) {
  datasets[[i]] <- rnorm(n, mean = mu, sd = sigma) 
}
```

For each dataset, save μ^ and the p-value arising from a test of H:μ=0 using α=0.05. Hint: to obtain the estimate and p-value, use broom::tidy to clean the output of t.test.
```{r}
alpha = 0.05

results <- data.frame(mu_h = numeric(num_datasets), p_value = numeric(num_datasets))


for (i in 1:num_datasets) {
  datasets[[i]] <- rnorm(n, mean = mu, sd = sigma)
  t_result <- broom::tidy(t.test(datasets[[i]], mu = 0))
  
  results$mu_h[i] <- t_result$estimate
  results$p_value[i] <- t_result$p.value
}

head(results)
```

Repeat the above for μ={1,2,3,4,5,6}, and complete the following:

Make a plot showing the proportion of times the null was rejected (the power of the test) on the y axis and the true value of μ
 on the x axis. Describe the association between effect size and power.
 
```{r}
mus = 1:6


set.seed(123)

power_results = data.frame(mu = mus, power = numeric(length(mus)))

for (i in seq_along(mu_values)) {
  mu = mus[i]
  reject_null = 0
  
  for (j in 1:num_datasets) {
    dataset = rnorm(n, mean = mu, sd = sigma)
    
    t_result = broom::tidy(t.test(dataset, mu = 0))
    p_value = t_result$p.value
    
    if (p_value < 0.05) {
      reject_null = reject_null + 1
    }
  }
  
  power_results$power[i] <- reject_null / num_datasets
}


```

```{r}
ggplot(power_results, aes(x = mu, y = power)) +
  geom_line(color = "blue") +
  geom_point(color = "blue") +
  labs(
    title = "power vs μ",
    x = "μ",
    y = "power"
  ) +
  theme_minimal()
```


Make a plot showing the average estimate of μ^ on the y axis and the true value of μ on the x axis. Make a second plot (or overlay on the first) the average estimate of μ^
 only in samples for which the null was rejected on the y axis and the true value of μ
 on the x axis. Is the sample average of μ^
 across tests for which the null is rejected approximately equal to the true value of μ
? Why or why not?
```{r}
estimate_results = data.frame(
  mu = mus,
  avg_all = numeric(length(mus)),
  avg_rejected = numeric(length(mus))
)



for (j in seq_along(mus)) {
  mu = mus[j]
  muh_all = numeric(num_datasets)
  muh_rejected = numeric(0)  # Only store mu_hat when null is rejected
  
  for (i in 1:num_datasets) {
    dataset = rnorm(n, mean = mu, sd = sigma)
    t_result = broom::tidy(t.test(dataset, mu = 0))
    muh = t_result$estimate
    p_value = t_result$p.value
    
    # Store the sample mean
    muh_all[i] = muh
    
    # If the null hypothesis is rejected, store mu_hat in the rejected set
    if (p_value < 0.05) {
      muh_rejected = c(muh_rejected, muh)
    }
  }
  
  # Calculate average mu_hat for all samples and for rejected samples only
  estimate_results$avg_all[j] = mean(muh_all)
  estimate_results$avg_rejected[j] = mean(muh_rejected, na.rm = TRUE)
}

```

```{r}
ggplot(estimate_results, aes(x = mu)) +
  geom_line(aes(y = avg_all, color = "Average μ^ (all samples)")) +
  geom_point(aes(y = avg_all, color = "Average μ^ (all samples)")) +
  geom_line(aes(y = avg_rejected, color = "Average μ^ (rejected only)")) +
  geom_point(aes(y = avg_rejected, color = "Average μ^ (rejected only)")) +
  labs(
    title = "Average Sample Estimate of μ^",
    x = "True Value of μ",
    y = "Average Estimate of μ^"
  ) +
  scale_color_manual(values = c("blue", "red"), name = "Estimates") +
  theme_minimal()
```




### Problem 3

Describe the raw data. 



Create a city_state variable (e.g. “Baltimore, MD”) and then summarize within cities to obtain the total number of homicides and the number of unsolved homicides (those for which the disposition is “Closed without arrest” or “Open/No arrest”).

```{r}
homicide_df = read_csv("./homicide-data.csv")|>
  mutate(city_state = paste (city, state, sep = ","))

citystate_df = homicide_df|>
  group_by(city_state)|>
  summarise(
    total_hom = n(),
    unsolved_hom = sum(disposition == "Closed without arrest" | 
                               disposition == "Open/No arrest", na.rm = TRUE)
  )
```
For the city of Baltimore, MD, use the prop.test function to estimate the proportion of homicides that are unsolved; save the output of prop.test as an R object, apply the broom::tidy to this object and pull the estimated proportion and confidence intervals from the resulting tidy dataframe.

```{r}
baltimore_df = homicide_df |>
  filter (city == "Baltimore" & state == "MD")

baltimore_homicide = nrow (baltimore_df)
baltimore_unsolved =  sum(baltimore_df$disposition == "Closed without arrest" | 
                          baltimore_df$disposition == "Open/No arrest", 
                          na.rm = TRUE)

```


```{r}
prop_result = tidy(prop.test(baltimore_unsolved, baltimore_homicide))
prop_esti = prop_result$estimate
ci = prop_result|>
  select (conf.low , conf.high)
```
Now run prop.test for each of the cities in your dataset, and extract both the proportion of unsolved homicides and the confidence interval for each. Do this within a “tidy” pipeline, making use of purrr::map, purrr::map2, list columns and unnest as necessary to create a tidy dataframe with estimated proportions and CIs for each city.
```{r}
city_result = citystate_df|>
  mutate (
    prop_test = map2(unsolved_hom, total_hom, ~ prop.test(.x,.y)),
    tidy_res = map(prop_test, tidy)
  )|>
  unnest (tidy_res)|>
  select (city_state, total_hom, unsolved_hom, estimate, conf.low, conf.high)
```
Create a plot that shows the estimates and CIs for each city – check out geom_errorbar for a way to add error bars based on the upper and lower limits. Organize cities according to the proportion of unsolved homicides.
```{r}
city_result = city_result|>
  arrange (desc(estimate))|>
  mutate(city_state = factor(city_state, levels = city_state))

ggplot(city_result, aes(x = city_state, y = estimate)) +
  geom_point(size = 1, color = "red") + 
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2, color = "darkgray") + 
  coord_flip() +  
  labs(
    title = "Proportion of Unsolved Homicides by City",
    x = "City",
    y = "Estimated Proportion of Unsolved Homicides"
  ) +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 8),
    plot.title = element_text(hjust = 0.5)
  )
```



















