---
title: "hw5"
author: "Senna"
date: "2024-11-05"
output: html_document
---


```{r}
library(tidyverse)
```

### Problem 1

Suppose you put n people in a room, and want to know the probability that at least two people share a birthday. For simplicity, we’ll assume there are no leap years (i.e. there are only 365 days) and that birthdays are uniformly distributed over the year (which is actually not the case).

Write a function that, for a fixed group size, randomly draws “birthdays” for each person; checks whether there are duplicate birthdays in the group; and returns TRUE or FALSE based on the result.

Next, run this function 10000 times for each group size between 2 and 50. For each group size, compute the probability that at least two people in the group will share a birthday by averaging across the 10000 simulation runs. Make a plot showing the probability as a function of group size, and comment on your results

```{r}

set.seed(1)

has_duplicate_birthday = function(n) {
  birthdays = sample(1:365, n, replace = TRUE)
  return(length(birthdays) != length(unique(birthdays)))
} #FALSE means there are no duplicate. TRUE means there are duplicate. 


group_sizes = 2:50  
num_it = 10000

p = numeric(length(group_sizes))

for (i in seq_along(group_sizes)) {
  n = group_sizes[i]
  
  sim_results = replicate(num_it, has_duplicate_birthday(n))
  p[i] = mean(sim_results)
}



ggplot(results_df, aes(x = group_sizes, y = p)) +
  geom_line(color = "blue") +
  geom_point(color = "red") +
  labs(title = "Probability of Shared Birthday vs. Group Size",
       x = "Group Size",
       y = "Probability of Shared Birthday") +
  theme_minimal()

```

### Problem 2

When designing an experiment or analysis, a common question is whether it is likely that a true effect will be detected – put differently, whether a false null hypothesis will be rejected. The probability that a false null hypothesis is rejected is referred to as power, and it depends on several factors, including: the sample size; the effect size; and the error variance. In this problem, you will conduct a simulation to explore power in a one-sample t-test.

First set the following design elements:
Fix n=30
Fix σ=5
Set μ=0
Generate 5000 datasets from the model x∼Normal[μ,σ]

```{r}
n = 30
sigma = 5
mu = 0
num_datasets = 5000

set.seed(123)  
datasets <- vector("list", num_datasets) 

for (i in 1:num_datasets) {
  datasets[[i]] <- rnorm(n, mean = mu, sd = sigma) 
}
```

For each dataset, save μ^ and the p-value arising from a test of H:μ=0 using α=0.05. Hint: to obtain the estimate and p-value, use broom::tidy to clean the output of t.test.
```{r}
alpha = 0.05

estimates = numeric(num_datasets)
p_values = numeric(num_datasets)

for (i in 1:num_datasets){
  t_results = broom::tidy(t.test(datasets[[i]], mu = 0))
  estimates[i] = t_results $ estimate
  p_values[i] = t_results $ p.value
  
}
```

Repeat the above for μ={1,2,3,4,5,6}, and complete the following:

Make a plot showing the proportion of times the null was rejected (the power of the test) on the y axis and the true value of μ
 on the x axis. Describe the association between effect size and power.
Make a plot showing the average estimate of μ^
 on the y axis and the true value of μ
 on the x axis. Make a second plot (or overlay on the first) the average estimate of μ^
 only in samples for which the null was rejected on the y axis and the true value of μ
 on the x axis. Is the sample average of μ^
 across tests for which the null is rejected approximately equal to the true value of μ
? Why or why not?
```{r}

```


1. power vs true mean
The first plot shows the relationship between the true mean and the power of the test. As the effect size (the distance of 
𝜇
μ from 0) increases, the power of the test also increases. This makes sense, as larger effect sizes are easier to detect with greater statistical power.
```{r}

```

2. average estmate of mu hat vs true mean (mu)
```{r}

```

